\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2014-05-19 paper.tex
Finished rebuttal letter + last fixes in the paper

\usepackage[draft, nomargin, marginclue, footnote]{fixme}

@@

We do not claim to address here the whole issue. This article attempts however to organise
it into a coherent challenge for artificial intelligence, and to explain and
illustrate some of the paths that we have investigated on our robots and that
result in a deliberative, knowledge-oriented, architecture for human-robot
interaction.

@@

Numerous approaches exist, like amodal \emph{proxies}~\cite{Jacobsson2008},
grounded amodal representations~\cite{Mavridis2006}, semantic
maps~\cite{Nuechter2008, Galindo2008,Blodow2011} or affordance-based planning
and object classification~\cite{Lorken2008, Varadarajan2011}.

@@

\fxfatal{Rephrase or remove}
It is sometimes difficult or even impossible to see and/or track an object in
certain states. This happens, for instance, when the object is hidden in a
container, when the robot holds it in its gripper, and more generally in any
state in which it is hidden by something else. {\sc Spark} models the possible
symbolic states of objects (whether the object is on a furniture, in an agent
hand, in a container, etc.): according to the robot perception of what has
happened since the object was last seen, the robot builds a belief of the
current possible state with associated probability model.

@@

Monitoring human activity is crucial for maintaining a coherent state of the
world. While full human action and activity recognition is a task that requires
knowledge and reasoning both on high level facts like goals, intentions and
plans, as well as bottom-up data from agent and object motions, simple temporal
and geometric reasoning on human hand trajectories and potential objects
placements can provide clues for high level human monitoring
processes. We call this temporal and geometric reasoning \emph{primitive action
recognition}. Those primitive actions are assessed through monitoring situations
like ``an empty hand is close to an object on a table'' (precursor for a
\emph{pick}), or ``a hand holding an object is over a container'' (precursor for
a \emph{throw}). {\sc Spark} recognises a handful \fxfatal{which ones?} of these
primitives, that are relied on by the robot to track the state transitions in
the human plans.

@@

Because HATP is a generic symbolic task planner, we have been able to design a
planning domain at a semantic level which is close to the one used in the
human-robot dialogue (the planner vocabulary contains concepts like
\texttt{give}, \texttt{table}, \texttt{is on}...). Hence only a few ontology
rules have been required to map both the knowledge extracted from the situation
assessment and the statements originated from the verbal interaction to the
planner domain.

Also, we would like to mention that, after some research (see appendix B
of~\cite{Lemaignan2012a} for a detailed discussion)  we have decided to
represent neither the planning domain nor the resulting plans in the knowledge
base: the planning domain (with task pre- and postconditions) is stored in a
specific format, outside of the central declarative knowledge repository, and
the plans are directly communicated to the robot controller. This leads to
independent declarative and procedural knowledge stores.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Robot Execution Control}
\label{sect|ctrl}

>>>>

\usepackage[nomargin, marginclue, footnote]{fixme}

@@

We do not claim to address here the issue as a whole. This article attempts
however to organise it into a coherent challenge for artificial intelligence,
and to explain and illustrate some of the paths that we have investigated on our
robots, that result in a set of deliberative, knowledge-oriented, software
components designed for human-robot interaction.

@@

Numerous approaches exist, like amodal (in the sense of modality-independent)
\emph{proxies}~\cite{Jacobsson2008}, grounded amodal
representations~\cite{Mavridis2006}, semantic
maps~\cite{Nuechter2008, Galindo2008,Blodow2011} or affordance-based planning
and object classification~\cite{Lorken2008, Varadarajan2011}.

@@

It is sometimes difficult or even impossible to see and/or track an object in
certain states. This happens, for instance, when the object is hidden in a
container, when the robot holds it in its gripper, and more generally in any
state in which it is hidden by something else. {\sc Spark} models the possible
symbolic states of objects (whether the object is on a furniture, in an agent
hand, in a container, etc.), and according to the robot perception of what has
happened since the object was last seen, the robot builds a belief of the
current possible state with associated probability model.

@@

Monitoring human activity is crucial for maintaining a coherent state of the
world. While full human action and activity recognition is a task that requires
knowledge and reasoning both on high level facts like goals, intentions and
plans, as well as bottom-up data from agent and object motions, simple temporal
and geometric reasoning on human hand trajectories and potential objects
placements can provide clues for high level human monitoring processes. We call
this temporal and geometric reasoning \emph{primitive action recognition}. Those
primitive actions are assessed through monitoring situations like ``an empty
hand is close to an object on a table'' (precursor for a \emph{pick}), or ``a
hand holding an object is over a container'' (precursor for a \emph{throw}).
{\sc Spark} recognises a core set of such primitives that are relied upon by the
robot to track the state transitions in the human plans.

@@

Because HATP is a generic symbolic task planner, we have been able to design a
planning domain at a semantic level which is close to the one used in the
human-robot dialogue (the planner vocabulary contains concepts like
\texttt{give}, \texttt{table}, \texttt{is on}...). This leads to a natural
mapping (low ``impedance mismatch'') between the knowledge extracted from the situation
assessment or the dialogue, and the planner.

Also, we would like to mention that, after some research we have decided against
representing the planning domain (\ie, the set of tasks with their pre- and
post-conditions) in the knowledge base itself due to expressiveness issues (see
appendix B of~\cite{Lemaignan2012a} for a detailed discussion). This effectively
leads to independent declarative and procedural knowledge stores.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Robot Execution Control}
\label{sect|ctrl}
